---
title: "Homework_1_Lape"
author: "Mike Lape"
date: "August 31, 2017"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
&nbsp;

Load the LearnBayes library.
```{r}
library(LearnBayes)
```

&nbsp;

1.	What is the size of the data? 
```{r}
# It has 657 rows (students) and 11 columns (variables).
dim(studentdata)
```

&nbsp;

2.	Show the top six rows of the data. 
```{r}
head(studentdata, 6)
```

&nbsp;

3.	Pull out the complete data of Students 4, 22, 517, and 533.
```{r}
extract <- studentdata[c(4,22,517,533),]
extract
```

&nbsp;

4.	Using the documentation command, describe each variable in the data. 
```{r,eval=FALSE}
help(studentdata, package = "LearnBayes")
```
Variable | Description from Docs
------------|------------------
Student | Student Number
Height | Height in Inches
Gender | Gender
Shoes | Number of Pairs of Shows Owned
Number | Number Chosen Between 1 and 10
DVDs | Name of Movie DVDs Owned
ToSleep | Time the Person Went to Sleep the Previous Night (Hours Past Midnight)
Haircut | Cost of Last Haircut Including Tip
Job | Number of Hours Working on a Job per Week
Drink | Usual Drink at Suppertime among Milk, Water, and Pop   

&nbsp;

5.	What is the nature of each variable? 
```{r}
sapply(studentdata, class)
```

&nbsp;

6.	Show the summary statistics of each variable.  
```{r}
summary(studentdata)
```

&nbsp;

7.	What is the gender distribution? 
```{r}
summary(studentdata$Gender)
```

&nbsp;

8.	What is the most favorite number of the students?  
```{r}
# It is 7.
# 191 students selected the number 7.
table(studentdata$Number)
```

&nbsp;

9.	What is the second most favorite number of the students? 
```{r}
# It is 5.
# 78 students chose the number 5.
table(studentdata$Number)
```

&nbsp;

10.	What is the least favorite number of students? 
```{r}
# It is 10. 
# Only 7 students selected the number 10.
table(studentdata$Number)
```

&nbsp;


11.	Use the 'table' command on 'studentdata$Shoes' and show the output. 
```{r}
table(studentdata$Shoes)
```

&nbsp;

12.	What is unusual about the output in question 11?  

Some students report having half pairs of shoes, this could indicate that they lost one of the shoes from that pair, although this is odd.  These students also report having a lot of pairs of shoes, some report 65 or 70 pairs or more.

&nbsp;

13.	Use the 'table' command on 'studentdata$Drink.'
```{r}
table(studentdata$Drink)
```
&nbsp;

14.	Cross-tabulate 'Gender' and 'Drink.'
```{r}
gender_drink_cross <- table(studentdata$Gender, studentdata$Drink)
gender_drink_cross
```

&nbsp;

15.	Calculate proportions row-wise and column-wise as well correct to two decimal places. 
```{r}
# Row-Wise:
round(prop.table(gender_drink_cross,1),2)
# Column-Wise:
round(prop.table(gender_drink_cross,2),2)
```