---
title: "Homework_4_Lape"
author: "Mike Lape"
date: "October 26, 2017"
output:
  word_document: default
---
&nbsp;

1.	Create function to estimate size of rhombus, based on surrounding sqaure.
```{r, fig.width=8, fig.height=8}

library(ggplot2)

# Long Diag = 4
# Short Diag = 2
SHORT = 2
LONG = 4

# Define R as long diag / 2
R = 4/2

rhom_area <- function(n)
{
  # Think of rhombus in square
  # Get the area of the square, and multiply by fraction of points that
  # fall within rhombus from all points that fall into square.

    
  # Generate our x,y df between -2 and 2 to cover square surrounding
  # this rhombus.  Generate rows equal to the handed in sample size, n.
  df = data.frame(x = runif(n, -2, 2), y = runif(n, -2, 2))
  
  # Determine if a point falls within rhombus
  # (2 * abs(df$x) + 1 * abs(df$y)) < R
  # 
  # Sum up the number of points in our df that fall within rhombus
  # sum(2*abs(df$x) + 1*abs(df$y) < R))
  #
  # Divide by n to get the ratio of points within rhombus to total points,
  # which is our ratio to transform area of square to area of rhombus.
  thetaHat = ((sum(2*abs(df$x) + 1*abs(df$y) < R)) / n)
  
  # Calculate that area
  areaHat = thetaHat * (2 * SHORT)^2
  
  return(list(areaHat = areaHat, N = n))
}
```

&nbsp;

2. Apply your rhom_area to 1:10000 and plot the results.
```{r, fig.width=8, fig.height=8}
# Run 10,000 simulations and plot the calculated area against the value of n
# to show the convergence on a true area of this rhombus.
#
# Comments:
#     When running this simulation 10k time and then plotting these results it
#     does indeed appear to converge upon the true area defined as:
#     (short leg * long leg) / 2
#     in this case:
#     (2 * 4) / 2 = 4
#

out = sapply(1:10000, rhom_area)
trueArea = (LONG * SHORT) / 2

p1 = ggplot(data = data.frame(area = unlist(out[1,]), n = unlist(out[2,])),
            aes(x = n, y = area)) + ggtitle("Convergence Upon True Area of Rhombus") +
            geom_point() + geom_hline(yintercept = trueArea, size=2, color = "red")
p1
```

&nbsp;

3. Write function that generates points inside the rhombus
```{r, fig.width=8, fig.height=8}
# This function will return a uniform distribution of points that
# fall inside the rhombus as pictured above in problem 1.
# The default n is 1, meaning it will return 1 x,y pair in 
# a dataframe that falls inside the rhombus, but you can
# pass it the number of pairs of x,y values you want and
# it will generate that many value pairs, and hand them back
# in a dataframe.
in_rhom <- function(n = 1)
{
  arr.x = rep(NA, n)
  arr.y = rep(NA, n)
  i = 0
  while(i < n)
  {
    x = runif(1, -2, 2)
    y = runif(1, -2, 2)
    
    if(((2 * abs(x)) + 1 * abs(y)) < R)
    {
      i = i + 1
      arr.x[i] = x
      arr.y[i] = y
    }
  }
  
  return(data.frame(x = arr.x, y = arr.y))
} 
```

4. Use rejection sampling to generate 1000 samples from the distribution below.
```{r, fig.width=8, fig.height=8}
# Distribution function provided
f <- function(x)
{
  dbeta(x, shape1 = 2, shape2 = 5)
}

# Simple distribution function
g <- function(x)
{
  dunif(x, 0.00, 1.00)
}

ratio.f.g = function(x)
{
  f(x)/g(x)
}

M = optimize(ratio.f.g, c(0,1), maximum = TRUE)
M = M$objective

genSamples <- function(n = 1000)
{
  # Create our sample vector and fill with NAs
  my.sample = rep(NA, n)
  
  i = 0
  j = 0
  while(i < n)
  {
    # Generate 1 value between 0 and 1
    y = runif(1, 0, 1)
    p = f(y) / ( M * g(y))
    accept = sample(c(TRUE,FALSE), 1, prob = c(p, 1-p))
    
    if(accept)
    {
      # Accepted number, add to vector
      my.sample[i+1] = y
      i = i + 1
    }
    else
    {
      j = j +1
    }
    
  }
  
  return(my.sample)
}
genSamples(1000)

```

5.  I used a simple proposal function of just dunif(x, 0.00, 1.00).  Running this
    function a few times with an n of 1,000, meaning we need to run this enough
    times to get 1000 accepted values, also produced ~1440 rejected values. So, the 
    acceptance ratio of this algorithm was was 0.694, or 69.4%
    
6.  Create a histogram with relative frequencies for your sample and add the true density curve to your histogram. 

```{r, fig.width=8, fig.height=8}
# Generate our samples
tmp = genSamples(1000)
# Create histogram
hist(tmp, freq = F, col = "grey", breaks = 100, xlim = c(0,1), main = "Histogram of Samples",
     ylab = "Density", xlab = "Samples")
curve(f(x),0, 1, add =T, col = "red")



```
