---
title: "Homework_6_Lape"
author: "Mike Lape"
date: "November 28, 2017"
output:
  word_document: default
---
&nbsp;

1. Find the bootstrap MSE and Margin of Error for the mean of the data (3 points).
```{r, fig.width=8, fig.height=8}
x = scan("rabbitblood.txt")
n = length(x)
boot = sample(x, n, replace = TRUE)

rep = 100000

thetahat = mean(x)
thetahatboot = rep(0,rep)

# Run the bootstrap rep times and store mean in i of thetahatboot
for(i in 1:rep)
{
  boot = sample(x, n, replace=TRUE); 
  thetahatboot[i] = mean(boot) 
}

# Calculate MSE
MSE = sum((thetahatboot - thetahat)^2) / rep
cat(paste0("\nMSE: ", round(MSE, 2)))

# Margin of Error is just 2 * Sqrt(MSE)
MoE = 2 * sqrt(MSE)
cat(paste0("\nMoE: ", round(MoE, 2)))
```

&nbsp;

2.	Write a function similar to BootstrapMSEmean to find the bootstrap bias for the mean of the data
```{r, fig.width=8, fig.height=8}
boot_bias = function(x, reps)
{
  n = length(x)
  
  # True Value
  thetahat = mean(x)
  thetahatboot = rep(0, reps)
  
  for(i in 1:reps)
  {
    boot = sample(x, n, replace=TRUE)
    thetahatboot[i] = mean(boot) 
  }
  
  bias = mean(thetahatboot) - thetahat
  cat(paste0("\nBias: ", round(bias, 2)))
}

# Call our function, which will handle the printing.
boot_bias(x, 1000)

```

&nbsp;


3.	Write a function similar to BootstrapMSEmean to find the MSE and Margin of Error for the median. (For a vector a, its median can be obtained by the command median(a)) .
```{r, fig.width=8, fig.height=8}
boot_med = function(x, reps)
{
  n = length(x)
  
  # True Value
  thetahat = median(x)
  thetahatboot = rep(0, reps)
  
  for(i in 1:reps)
  {
    boot = sample(x, n, replace=TRUE)
    thetahatboot[i] = median(boot) 
  }
  
  
  # Calculate and print MSE
  MSE = sum((thetahatboot - thetahat)^2) / rep
  cat(paste0("\nMSE: ", round(MSE, 2)))
  
  # Margin of Error is just 2 * Sqrt(MSE)
  MoE = 2 * sqrt(MSE)
  cat(paste0("\nMoE: ", round(MoE, 2)))
}
# Call our function, which will handle the printing.
boot_med(x, 1000)

```

&nbsp;
4. Find the 90% bootstrap confidence interval for μ, the population mean
```{r, fig.width=8, fig.height=8}
reps = 10000

# Mean and Std Dev for actual X
meanx = mean(x)
sdx = sd(x)

n = length(x)
var = rep(0,reps)

for(i in 1:reps)
{
  # Run the bootstrap and calculate boot mean and std dev
  boot = sample(x, n, replace=TRUE) 
  boot_mean = mean(boot)
  boot_sd = sd(boot)
  
  # Calculate the variance and shove it in array        
  var[i] = (boot_mean - meanx) / (boot_sd / sqrt(n))
}

alpha = .1
lower = quantile(var, alpha/2)
upper = quantile(var, 1-alpha/2)

left = meanx - upper*sdx / sqrt(n)
right= meanx - lower*sdx / sqrt(n)
c(left,right)

```

&nbsp;

5. 	Find the 95% bootstrap confidence interval for σ^2, the population variance
```{r, fig.width=8, fig.height=8}
# Similar to #4 just swapping out mean and swapping in variance
reps = 10000

# Mean and Std Dev for actual X
meanx = mean(x)
sdx = sd(x)

n = length(x)
var = rep(0,reps)

for(i in 1:reps)
{
  # Run the bootstrap and calculate boot mean and std dev
  boot = sample(x, n, replace=TRUE) 
  boot_mean = mean(boot)
  boot_sd = sd(boot)
  
  # Calculate the variance and shove it in array        
  var[i] = ( (n - 1) * (boot_sd)^2)/(sdx)^2
}

alpha = .05
lower = quantile(var, alpha/2)
upper = quantile(var, 1-alpha/2)

left = (n - 1)*sdx^2 /upper
right = (n - 1)*sdx^2 /lower
c(left,right)

```

&nbsp;

6. Find the 95% bootstrap confidence interval for ρ, the population correlation between x and y 
```{r, fig.width=8, fig.height=8}
y = read.table("height.txt")$V1
x = read.table("height.txt")$V2

n = length(x)
m = length(y)

# Grab indices for bootstrap then put values in correct vector
boot_index = sample(1:n, n, replace=TRUE)
xboot = x[boot_index]         
yboot = y[boot_index]         


reps = 1000

thetahat = cor(x,y)
thetahatboot = rep(0,reps)

for(i in 1:reps)
{
  
  # Grab indices for bootstrap then put values in correct vector
  boot_index=sample(1:n, n, replace=TRUE)
  xboot = x[boot_index]        
  yboot = y[boot_index]     
    
  if((var(xboot)!=0)&(var(yboot)!=0))
  {
    thetahatboot[i]=cor(xboot, yboot)#sample corr for bootstrap sample
  }
}

quantile(thetahatboot, prob = c(0.025, 0.975), na.rm = TRUE)
```


&nbsp;

7. Assume normality. Find a 90% confidence interval for β_1
```{r, fig.width=8, fig.height=8}
fit = lm(y~x)

alpha = 0.1
e = fit$residuals
sx = sd(x)
n = length(x)
SE = sqrt(sum(e^2)/((n-2)*(n-1)*(sx^2)))
left = fit$coef[2]-SE*qt(1-alpha/2,n-2)
right = fit$coef[2]-SE*qt(alpha/2,n-2)
c(left,right)


```

&nbsp;
8.	Describe the difference between Methods 1 and 2 for bootstrap for regression. 
```{r, fig.width=8, fig.height=8}
# Method 1 is used for x,y pairs where both the x and y are random.    
# Method 2 is used for when just the y is random, and the x is not.
```

&nbsp;

9.	Obtain the estimates β ^_0 and β ^_1 
```{r, fig.width=8, fig.height=8}

fit = lm(y ~ x)
fit$coef[2]

reps = 1000


b0 = rep(0, reps)
b1 = rep(0, reps)

for(i in 1:reps)
{
  boot_index=sample(1:n, n, replace=TRUE)
  xboot = x[boot_index]        
  yboot = y[boot_index]     
  
  # Store the 
  fit_boot = lm(yboot ~ xboot)
  b0[i] = fit_boot$coef[1]
  b1[i] = fit_boot$coef[2]
    
}

est_b_0 = mean(b0)
est_b_1 = mean(b1)

cat(paste0("\n", reps, " reps estimate of B0: ", round(est_b_0, 2)))
cat(paste0("\n", reps, " reps estimate of B1: ", round(est_b_1, 2)))
```

&nbsp;

10. Calculate the 95% bootstrap confidence interval for β_1 (assuming that X is fixed)

```{r, fig.width=8, fig.height=8, eval = FALSE}

SE_beta1 = function(x, y)
{
  fit = lm(y ~ x)
  e = fit$residuals
  sd_x = sd(x)
  n = length(x)
  sqrt(sum(e^2)/((n-2)*(n-1)*(sd_x^2)))
}



fit = lm(y ~ x)
e = fit$residuals
sd_x = sd(x)
t_boot = rep(0,reps)

reps = 1000

for(i in 1:reps)
{
  e_boot = sample(e, replace=TRUE)
  y_boot = fit$coef[1] + fit$coef[2]*x + e_boot
  
    # now I have a bootstrap sample (x, yb)
  fit_boot = lm(y_boot ~ x);
  t_boot[i]=(fit_boot$coef[2] - fit$coef[2])/SE_beta1(x,y_boot)
}

# 95% CI
alpha = 0.05

left = fit$coef[2]-SE_beta1(x,y)*quantile(t_boot,1-alpha/2)
right = fit$coef[2]-SE_beta1(x,y)*quantile(t_boot,alpha/2)


c(left,right)



```